---
title: "Rule Trends"
author: "Adam Ribaudo"
date: "3/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(arules)
library(tidyverse)
library(data.table)
library(readxl)
library(knitr)
library(arulesViz)
library(PropCIs)
library(ggthemes)
library(scales)
require(extrafont)
library(RColorBrewer)
#library(usedist)
source("functions.R")

my_theme <- theme(
        text=element_text(size=14,  family="Open Sans"),
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.background  = element_blank(),
        plot.background = element_rect(fill="white", colour=NA),
        panel.grid = element_line(color = "grey90"),
        plot.tag.position = "top",
        plot.margin =  unit(c(5,5,5,5), "pt"),
        plot.title = element_text(size = unit(c(14),"pt"),
                                  margin = unit(c(10,0,0,0),"pt"),
                                  hjust = 0.5),
        plot.subtitle = element_text(size = unit(c(10),"pt"),
                                  margin = unit(c(8,0,0,0),"pt"),
                                  hjust = 0.5)
        ) 
```

# Load data

```{r warning=FALSE,message=FALSE}
transaction_col = "transaction"
partition_col = "partition"
item_col = "item"
weight_col = "weight"
minsup <- .03
minconf <- .5 #TODO would like to work with higher levels of confidence, but need to lower support which causes the computation to take too long

critical_z <- -1.65 # equates to 95% confidence
confidence_val <- .95
partition_levels <- month.abb

# Functions Require that the df have partition, transaction, and item as columns and that the parameters above are set
df <- read_excel("..\\data\\online_retail_II.xlsx", sheet=1) %>% bind_rows(read_excel("..\\data\\online_retail_II.xlsx", sheet=2)) %>%
  rename(InvoiceNo = Invoice, UnitPrice = Price, CustomerID = `Customer ID`) %>%
  mutate(isRefund = grepl("C", InvoiceNo) | Quantity < 0) %>%
  filter(isRefund == FALSE) %>%
  # Convert datetime to date
  mutate(InvoiceDate = as.Date(InvoiceDate, format = "%m/%d/%Y %H:%M")) %>%
  # Extract month
  mutate(InvoiceYearMonth = paste0(year(InvoiceDate),month(InvoiceDate))) %>%
  # Consider 2010 data only
  filter(grepl("2010",InvoiceYearMonth)) %>%
  mutate(month = month.abb[month(InvoiceDate)]) %>%
  mutate(revenue = Quantity * UnitPrice) %>%
  filter(!grepl("TEST", StockCode), revenue > 0, !is.na(CustomerID)) # remove revenue of 0 and missing customer Ids

df_par <- df %>% rename(partition = month, transaction = InvoiceNo, item = StockCode, weight_abs = revenue)

# having trouble running at .01 support
results <- get_change_mining_results(df_par,minsup=minsup,minconf=minconf,confidence_val=confidence_val, useWeights = F)
```

# Evaluation of All 2010 Data

## Number of Frequent Itemsets

```{r message=FALSE, warning=FALSE}
supports <- c(.015, .02, .025, .03, .035)

global_support_df <- data.frame(support = supports)
global_support_df$frequent_items <- as.vector(sapply(supports,function(x)length(eclat(get_transactions_from_df(df_par), parameter = list(support = x, minlen=1), control=list(verbose=F)))))

#png("freq_itemsets_by_support.png", units="in", width=8, height=5, res=300)
ggplot(global_support_df) + geom_col(aes(x=factor(support),y=frequent_items)) + labs(y="Frequent Itemsets",x="Support",title="Number of Frequent Itemsets by Min Support") + my_theme
#dev.off() 
```

## Frequent Itemsets by Month

```{r message=FALSE, warning=FALSE}
supports <- c(.02, .025, .03, .035)

supports_by_month_df <- bind_rows(lapply(partition_levels, function(month){
    cur_transactions <- partitioned_transactions[[month]]
  support_df <- data.frame(support = supports)
  support_df$month <- month
  
  support_df$frequent_items <- as.vector(sapply(supports,function(x)length(eclat(cur_transactions, parameter = list(support = x, minlen=1), control=list(verbose=F)))))
  support_df
}))

supports_by_month_df <- supports_by_month_df %>% mutate(month = factor(month, level = partition_levels))

#png("freq_itemsets_by_support_by_month.png", units="in", width=8, height=5, res=300)
ggplot(supports_by_month_df) + geom_col(aes(x=factor(support),y=frequent_items)) + facet_wrap(~month) + labs(y="Frequent Itemsets",x="Support",title="Number of Frequent Itemsets by Min Support and Month") + my_theme
#dev.off() 
```

## Rule Counts by Minsup

This chart shows the number of frequent rules, semi stable rules, and stable rules at varying levels of minsup. Ideally, we can set a minsup that yields a small handful of stable rules worth evaluating further.

This shows the value of finding stable rules: they help reduce the number of frequent rules we must consider

```{r message=FALSE, warning=FALSE}
# Plot # of rules by minsup
minsups <- c(.02,.025,.03,.035)
rule_type_counts_unweighted <- get_rule_type_counts(minsups, df_par, minconf, confidence_val, useWeights = F)

#png("rule_type_counts_by_support.png", units="in", width=8, height=5, res=300)
ggplot(rule_type_counts_unweighted, aes(x=minsup, y=`Rule Count`)) + geom_line(aes(color=`Rule Type`)) + labs(title="Rule Count as Minsup Increases, by Rule Type", subtitle="Invoices Partitioned by Month") + my_theme + scale_y_log10() + xlab("Minimum Support")
#dev.off() 
```


## Rules Scatterplot for All Rules Meeting Minsup in at least a Single Month

```{r warning=FALSE,message=FALSE}
#png("rules_scatterplot.png", units="in", width=8, height=5, res=300)
plot(results$semi_stable_rules, control = list(main = "Semi Stable Rules"))
#dev.off() 
```



```{r}
#png("rules_graph.png", units="in", width=8, height=5, res=300)
plot(results$semi_stable_rules, method="graph", control = list(main = "Semi Stable Rules Graph"))
saveAsGraph(results$semi_stable_rules, file = "semi_stable_rules.graphml")
#dev.off()
```

# Partition Data by Month

## Distance between Months

The heatmap below shows the "distance" between each month as calculated by the Jaccard distance: (1-intersection/union) where intersection and union are comparing frequent itemsets between the 2 months.

```{r echo=FALSE,warning=FALSE,message=FALSE}
distance_df <- as.data.frame(combn(partition_levels,2)) %>% 
  transpose() %>% 
  rename(month1=1,month2=2) %>% 
  mutate(month1=factor(month1,levels=partition_levels)) %>%
  mutate(month2=factor(month2,levels=partition_levels)) %>%
  rowwise %>%
  mutate(distance = get_partition_dist(results$partitioned_transactions[[as.character(month1)]],results$partitioned_transactions[[as.character(month2)]],minsup=minsup, useWeights = F, target = "frequent itemsets")) 

#png("month_distance.png", units="in", width=8, height=5, res=300)
ggplot(distance_df %>%
  bind_rows(data.frame(list(month1=partition_levels, month2=partition_levels, distance = NA))), aes(x=month1,y=month2,fill=distance)) + geom_tile() + my_theme + labs(title = "Jaccard Distance between Months", subtitle = "Minimum Support = .03", x="", y="") + scale_fill_distiller(direction=1)
#dev.off()
```

# Extension Sets

Example: {A,B} is found in Window 1 and {A,B,C} is found in Window 2 but not Window 1. sup W2 / sup W1 = inf

```{r include=FALSE, message=FALSE, warning = FALSE}
extensions_df <- expand.grid(partition_levels, partition_levels) %>% rename(month1 = Var1, month2 = Var2) %>% 
  filter(month1 != month2) %>%
  rowwise() %>% mutate(distance = get_partition_dist(results$partitioned_transactions[[as.character(month1)]],results$partitioned_transactions[[as.character(month2)]],minsup=minsup, useWeights = F, target = "frequent itemsets")) %>%
mutate(extension_sets = get_num_extension_sets(results$partitioned_transactions[[as.character(month1)]], results$partitioned_transactions[[as.character(month2)]]) ) #%>%
  #mutate(reduction_sets = get_num_extension_sets(results$partitioned_transactions[[as.character(month2)]], results$partitioned_transactions[[as.character(month1)]]) )

#png("extensions.png", units="in", width=8, height=5, res=300)
ggplot(extensions_df) + geom_point(aes(x = distance, y=extension_sets)) + 
  geom_text(nudge_y = .3, aes(x=distance, y=extension_sets, label = 
                                if_else(extension_sets > 3, paste(as.character(month1), "->",as.character(month2)), ""))) + 
  my_theme + labs(x="Jaccard Distance", y="Extension Sets", title = "Extension Sets Between Months") 
#dev.off()
```

## Example Extension Sets 

```{r}
get_extension_sets(results$partitioned_transactions[[as.character("Feb")]], results$partitioned_transactions[[as.character("Aug")]]) %>% rename(Feb = t0_items, Aug = t1_items) %>% gt
```

## Itemset size across Months

Each cell is a histogram showing the size of the frequent itemsets (with minsup=`r minsup`) for each month partition. The red line and coloring denote the median for that month. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

itemset_distribution_df <- as.data.frame(rbindlist(lapply(partitioned_transactions, function(x)data.frame(size=size(x))),fill=T,idcol=T)) %>% rename(partition = .id) %>%
   mutate(partition = factor(partition, levels=partition_levels)) %>%
  group_by(partition) %>% mutate(`Median` = median(size))

png("month_transaction_size_distribution.png", units="in", width=8, height=5, res=300)
ggplot(itemset_distribution_df) + geom_histogram(aes(x=size),bins=10) + 
  facet_wrap(~partition) + 
  geom_vline(aes(xintercept = `Median`),col='red',size=.1) +
  xlim(c(0,50)) + labs(x="Transactions Size",y="Number of Transactions",title="Transaction Size Distribution by Month", subtitle = "Red line denotes median size") + my_theme
dev.off()
```

## Frequent itemset size by month

```{r echo=FALSE, message=FALSE, warning=FALSE}

itemset_distribution_df <- as.data.frame(rbindlist(lapply(partitioned_transactions, function(x)data.frame(size=size(eclat(x, parameter = list(support = .03), control = list(verbose=F))))),fill=T,idcol=T)) %>% rename(partition = .id) %>%
   mutate(partition = factor(partition, levels=partition_levels)) %>%
  group_by(partition) %>% mutate(`Median` = median(size))

#png("month_transaction_size_distribution.png", units="in", width=8, height=5, res=300)
ggplot(itemset_distribution_df) + geom_histogram(aes(x=size),bins=10) + 
  facet_wrap(~partition) + 
  geom_vline(aes(xintercept = `Median`),col='red',size=.1) +
  xlim(c(0,5)) + labs(x="Transactions Size",y="Number of Transactions",title="Frequent Itemset Size Distribution by Month") + my_theme
#dev.off()
```

## Stable Rules

The table below shows rules that are stable across all 12 months at minsup = `r minsup`  

21977 = PACK OF 60 PINK PAISLEY CAKE CASES  
21212 = PACK OF 72 RETRO SPOT CAKE CASES

```{r echo=FALSE, message=FALSE, warning=FALSE}
kable(results$stable_df)
```

## Confidence Interval for stable rules

For the stable rules found, we can plot their support confidence intervals by month to show that they are, in fact, stable.

```{r echo=FALSE, message=FALSE, warning=FALSE}
ci_graph_df <- get_ci_graph_df(results$partitioned_rules_measures_df, results$stable_df, confidence_val, partition_levels)

png("stable_rules_scatterplot.png", units="in", width=8, height=2, res=300)
ggplot(ci_graph_df, aes(x=partition, y=support)) + 
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=.1) +
    #geom_line() +
    geom_point() + theme(legend.position="bottom") + facet_wrap(~rule,strip.position="top",nrow=2) + my_theme + 
  labs(title = "Stable Rules", x = "Month", y="Support")
dev.off()
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ci_graph_df <- get_ci_graph_df(results$partitioned_rules_measures_df, head(results$semi_stable_df,3), confidence_val, partition_levels)

#png("semi_stable_rules_ci.png", units="in", width=8, height=5, res=300)
ggplot(ci_graph_df, aes(x=partition, y=support)) + 
    geom_errorbar(aes(ymin=ci_low, ymax=ci_high), width=.1) +
    #geom_line() +
    geom_point() + theme(legend.position="bottom") + facet_wrap(~rule,strip.position="top",nrow=3) + my_theme + 
  labs(title = "Semi Stable Rules", x = "Month", y="Support")
#dev.off()
```


